Mechanization of Separation in Generic Extensions

### 2019-01-05 LICS/conclusion_m.tex
Profundizando un poco en la dicotomía internalizado/externalizado.

This is an appropriate spot to insist that the internal/external
dichotomy has been a powerful agent in the shaping of our project. It
traces back to the definition of the formula-transformer $\forceisa$,
which must be done by recursion, and this requirement spills
indirectly to the proof of the Separation Axiom (despite the latter is
not by induction). Now the need of using internalized formulas
reappears with the need of having a general version of the Reflection
Principle (because the $\Phi$ there depends on the instance of
Replacement being proved).

>>>>

This is an appropriate point to insist that the internal/external
dichotomy has been a powerful agent in the shaping of our project.
This tension was also pondered by Paulson in his formalization of
Gödel's constructible universe \cite{MR2051585}; after choosing a
shallow embedding of $\ZF$, every argument proved by induction on
formulas (or functions defined by recursion) should be done using
internalized formulas. Working on top of Paulson's library, we
prototyped the formula-transformer $\forceisa$ is defined for
internalized formulas and this affects indirectly the proof of the
Separation Axiom (despite the latter is not by induction). The need of
using internalized formulas reappears with the need of a general
version of the Reflection Principle (because the $\Phi$ there depends
on the instance of Replacement being proved).
